Constraints on Assembly Bias from Galaxy Clustering

### 2016-06-17 ms.tex
Removed reduncancy of summation sign.
Minor changes to equation.

In addition to the standard occupation statistics described in the
previous section, in this paper we also use the decorated HOD
formalism to connect galaxies to dark matter halos in a manner that
has simultaneous dependence on both $\mvir$ and {\em halo concentration}.
Briefly, we use Equations \ref{eq:ncen} and \ref{eq:nsat} as our
``baseline" first occupation moments.  At fixed $\mvir,$ halos are
divided into one of two categories, those of high- and
low-concentration, depending on whether the concentration of the halo
places it above or below the rank-order percentile $f_{\rm split},$
which we keep fixed to $f_{\rm split}=0.5$ throughout the paper for
simplicity. High-concentration halos have a different first
occupation moment relative to low-concentration halos of the same
mass,
%
\begin{eqnarray}
\langle\ngal|\mvir, c_{\rm high}\rangle & = & \langle\ngal|\mvir \rangle + \delta\ngal \nonumber \\  
\langle\ngal|\mvir, c_{\rm low}\rangle  & = & \langle\ngal|\mvir \rangle - \delta\ngal
\end{eqnarray}
%
This difference, $\delta\ngal$, between the
first moment of high- and low-concentration halos is modulated by
$\abias,$ the novel parameter of the decorated HOD governing assembly
bias. Values of $\abias=\pm1$ correspond to the maximum strength of
assembly bias allowable by the constraint that the model preserves the
marginalized first moment, $\langle\ngal|\mvir\rangle;$ thus
regardless of the value of $\abias,$ in the decorated HOD the
marginalized first moment of centrals and satellites are {\em
  unchanged from the baseline value defined by Equations \ref{eq:ncen}
  and \ref{eq:nsat}}. In other words, {\em decorated HOD models all have the same HODs,
when averaged over all halos at fixed mass, as standard HOD
models with the same five standard HOD parameters}. 
The only change in Decorated HOD models is whether or not an
additional property also modulates halo occupation at fixed halo
mass. A value of $\abias = 0$ indicates no galaxy assembly bias
whatsoever. We refer the reader to \citet{hearin_etal16} for further
details about the decorated HOD.

In the present work, we fix our model to the simplest class of galaxy
assembly bias models, though a recipe for generalizing to more
complicated models can be found in \citet{hearin_etal16}. In
particular, we split halos into two populations as specified in the
previous paragraph. We then populate halos with satellite galaxies
specified by an assembly bias parameter $-1 \le A_{\rm sat} \le 1$ and
central galaxies with a distinct assembly bias parameter
$-1 \le A_{\rm cen} \le 1$. For the value of $f_{\rm split}=0.5$ adopted here,
we then have that
%
\begin{eqnarray}
  \delta N_{\rm cen} & = & A_{\rm cen} \, {\rm MIN}\left\{\langle N_{\mathrm{cen}}|\mvir\rangle , 1- \langle N_{\mathrm{cen}}|\mvir\rangle\right\}\nonumber\\
  \delta N_{\rm sat} & = & A_{\rm sat} \, \langle N_{\mathrm{sat}}|\mvir\rangle
\end{eqnarray}
%
In our analyses that include assembly bias, we vary these two
additional parameters, bringing the total number of parameters
that vary in these analyses to seven. As we show below, these additional
parameters are in many instances poorly constrained by clustering data
of the quality of SDSS DR7 alone, so exploring more complex models of
assembly bias does not yet seem justified in such analyses.

@@

We constrain HOD parameters based on SDSS DR7 measurements of the projected
galaxy two-point functions, $w_{\rm p}(r_{\rm p})$, and galaxy number densities, $n_{\rm g}$,
for luminosity threshold samples published in \citet{zehavi_etal11}. We use the full covariance
matrix of the projected correlation function available at
{\tt astroweb.cwru.edu/izehavi/dr7\_covar/table8}. We assume a likelihood of the
form $\mathcal{L} \propto e^{-\chi^2/2}$, where
%
\begin{equation}
\label{eq:chisquare}
\chi^2 = \sum_{i,j} \Delta w_{{\rm p},i} \, [C^{-1}]_{ij} \, \Delta w_{{\rm p},j} + \frac{(n_{\rm g}^{\rm mock} - n_{\rm g}^{\rm meas})^2}{\sigma_{\rm n}^2},
\end{equation}
%
$\Delta w_{{\rm p},i} = w_{\rm p}^{\rm mock}(r_{{\rm p},i}) - w_{\rm p}^{\rm meas}(r_{{\rm p},i})$ is the difference
between the projected two-point function predicted by the mock catalog and the measured value in the i$^{\rm th}$
separation bin (of 12 bins total), $C^{-1}$ is the inverse of the covariance matrix of the measurements,
and the last term in Eq.~(\ref{eq:chisquare}) is the contribution from the difference between the predicted and measured
galaxy number density. In Eq.~(\ref{eq:chisquare}), the we adopt the Einstein summation convention and sum over the 
dummy indices $i$ and $j$. The error on the galaxy number density assumes Poisson statistics for both the measured and predicted galaxy number densities.

>>>>

In addition to the standard occupation statistics described in the
previous section, in this paper we also use the decorated HOD
formalism to connect galaxies to dark matter halos in a manner that
has simultaneous dependence on both $\mvir$ and {\em halo concentration}.
Briefly, we use Equations \ref{eq:ncen} and \ref{eq:nsat} as our
``baseline" first occupation moments.  At fixed $\mvir,$ halos are
divided into one of two categories, those of high- and
low-concentration, depending on whether the concentration of the halo
places it above or below the rank-order percentile $f_{\rm split},$
which we keep fixed to $f_{\rm split}=0.5$ throughout the paper for
simplicity. High-concentration halos have a different first
occupation moment relative to low-concentration halos of the same
mass,
%
\begin{eqnarray}
\langle\ngal|\mvir, c_{\rm high}\rangle & = & \langle\ngal|\mvir \rangle + \delta\ngal \nonumber \\  
\langle\ngal|\mvir, c_{\rm low}\rangle  & = & \langle\ngal|\mvir \rangle - \delta\ngal
\end{eqnarray}
%
This difference, $\delta\ngal$, between the
first moment of high- and low-concentration halos is modulated by
$\abias,$ the novel parameter of the decorated HOD governing assembly
bias. Values of $\abias=\pm1$ correspond to the maximum strength of
assembly bias allowable by the constraint that the model preserves the
marginalized first moment, $\langle\ngal|\mvir\rangle;$ thus
regardless of the value of $\abias,$ in the decorated HOD the
marginalized first moment of centrals and satellites are {\em 
unchanged from the baseline value defined by Equations (\ref{eq:ncen}) 
and (\ref{eq:nsat})}. In other words, {\em decorated HOD models all have the same HODs,
when averaged over all halos at fixed mass, as standard HOD
models with the same five standard HOD parameters}. 
The only change in Decorated HOD models is whether or not an
additional property also modulates halo occupation at fixed halo
mass. A value of $\abias = 0$ indicates no galaxy assembly bias
whatsoever. We refer the reader to \citet{hearin_etal16} for further
details about the decorated HOD.

In the present work, we fix our model to the simplest class of galaxy
assembly bias models, though a recipe for generalizing to more
complicated models can be found in \citet{hearin_etal16}. In
particular, we split halos into two populations as specified in the
previous paragraph. We then populate halos with satellite galaxies
specified by an assembly bias parameter $-1 \le A_{\rm sat} \le 1$ and
central galaxies with a distinct assembly bias parameter
$-1 \le A_{\rm cen} \le 1$. For the value of $f_{\rm split}=0.5$ adopted here,
we then have that
%
\begin{eqnarray}
%
\delta N_{\rm cen} & = & A_{\rm cen} \, {\rm MIN} \left[ \langle N_{\mathrm{cen}} \vert \mvir \rangle ,
1- \langle N_{\mathrm{cen}} \vert \mvir \rangle \right]  \nonumber \\ 
\delta N_{\rm sat} & = & A_{\rm sat} \, \langle N_{\mathrm{sat}} \vert \mvir \rangle
\end{eqnarray}
%
In our analyses that include assembly bias, we vary these two
additional parameters, bringing the total number of parameters
that vary in these analyses to seven. As we show below, these additional
parameters are in many instances poorly constrained by clustering data
of the quality of SDSS DR7 alone, so exploring more complex models of
assembly bias does not yet seem justified in such analyses.

@@

We constrain HOD parameters based on SDSS DR7 measurements of the projected
galaxy two-point functions, $w_{\rm p}(r_{\rm p})$, and galaxy number densities, $n_{\rm g}$,
for luminosity threshold samples published in \citet{zehavi_etal11}. We use the full covariance
matrix of the projected correlation function available at
{\tt astroweb.cwru.edu/izehavi/dr7\_covar/table8}. We assume a likelihood of the
form $\mathcal{L} \propto e^{-\chi^2/2}$, where
%
\begin{equation}
\label{eq:chisquare}
\chi^2 = \Delta w_{{\rm p},i} \, [C^{-1}]_{ij} \, \Delta w_{{\rm p},j} + \frac{(n_{\rm g}^{\rm mock} - n_{\rm g}^{\rm meas})^2}{\sigma_{\rm n}^2},
\end{equation}
%
$\Delta w_{{\rm p},i} = w_{\rm p}^{\rm mock}(r_{{\rm p},i}) - w_{\rm p}^{\rm meas}(r_{{\rm p},i})$ is the difference
between the projected two-point function predicted by the mock catalog and the measured value in the i$^{\rm th}$
separation bin (of 12 bins total), $C^{-1}$ is the inverse of the covariance matrix of the measurements,
and the last term in Eq.~(\ref{eq:chisquare}) is the contribution from the difference between the predicted and measured
galaxy number density. In Eq.~(\ref{eq:chisquare}), the we adopt the Einstein summation convention and sum over the 
dummy indices $i$ and $j$. The error on the galaxy number density assumes Poisson statistics for both the measured and predicted galaxy number densities.
